---
title: "Coherent streamflow variability in Monsoon Asia over the past eight centuries---links to oceanic drivers"
subtitle: "All time series plots"
author: "Hung Nguyen, Sean Turner, Brendan Buckley, and Stefano Galelli"
output: 
  html_document: 
    highlight: tango
    number_sections: yes
    theme: journal
    toc: yes
    toc_float: yes
    df_print: paged
---

# Introduction

This document provides all time series plots from the reconstructions. You should read this only after going through the main document (`paleo_asia.Rmd`). Be warned: lots of scrolling here.


The following code chunk sets up the environment and reads the data.

```{r setup, message=FALSE, warning=FALSE}
source('R/init.R')

# Read data
instQmeta <- fread('data/instQ_meta.csv', key = 'ID') # Streamflow metadata
instQ <- fread('data/instQ.csv', key = 'ID')          # Instrumental streamflow

ldsFit <- readRDS('results/ldsFit.RDS')
ldsRec <- lapplyrbind(ldsFit, '[[', 'rec', id = 'ID')
ldsRec2 <- lapplyrbind(ldsFit, '[[', 'rec2', id = 'ID')
setkey(ldsRec, ID)
setkey(ldsRec2, ID)
```


# All instrumetal period comparisons

Like **Figure S6**, but with all stations.

```{r inst plot, fig.width=8, fig.height=24}
DT <- merge(ldsRec2, instQ, by = c('ID', 'year'))
ggplot(DT) +
  geom_ribbon(aes(year, ymin = Ql, ymax = Qu, fill = '95% Confidence Interval'), 
              alpha = 0.25) +
  geom_line(aes(year, Q, colour = 'Reconstruction')) +
  geom_line(aes(year, Qa, colour = 'Observation')) +
  facet_wrap(vars(ID), ncol = 3, scales = 'free', labeller = as_labeller(ID_to_name_basin)) +
  scale_colour_manual(name = NULL, values = c('darkorange', 'black')) +
  scale_fill_manual(name = NULL, values = 'gray') +
  scale_x_continuous(breaks = seq(1960, 2010, 10)) +
  labs(x = NULL, y = 'Mean annual flow [m\u00b3/s]') +
  theme(legend.position = 'top',
        legend.key.width = unit(1.5, 'cm'))
```

# Full reconstructions

Time series for full reconstructions (like **Figure S7**, but for all stations).

```{r flow history time series, fig.width=8, fig.height=62}
lp <- copy(ldsRec)[, lp := dplR::pass.filt(Q, 20, 'low', 'Butterworth')]
ggplot() +
  geom_rect(aes(xmin = firstYear, xmax = finalYear, ymin = -Inf, ymax = Inf), megadroughts,
            fill = 'darkorange', alpha = 0.2) +
  geom_hline(aes(yintercept = Qm, colour = 'Long term mean'), 
             lp[, .(Qm = mean(Q)), by = ID]) +
  geom_line(aes(year, Q, colour = 'Reconstruction'), lp, size = 0.2) +
  geom_line(aes(year, lp, colour = '20-yr low pass'),
            lp[, .SD[30:(.N-30)], by = ID], size = 0.5) +
  facet_wrap(vars(ID), ncol = 1, scales = 'free_y', strip.position = 'right') +
  labs(x = NULL, y = 'Q [m\u00b3/s]') +
  scale_x_continuous(breaks = seq(1200, 2000, 50), expand = c(0, 10)) +
  scale_colour_manual(name = NULL, values = c('black', 'maroon', 'gray70')) +
  theme(legend.position = 'top', 
        legend.key.width = unit(2, 'cm'))
```

# Full ensembles

Now we plot the full ensemble for all sations (in gray) compared with the best member that we chose above (in blue).

```{r ensemble ts plot, fig.width=8, fig.height=62}
ensembleRec <- readRDS('results/ensemble_reconst.RDS')
ggplot(ensembleRec) +
  geom_line(aes(year, Q, group = interaction(kwf, p)), colour = 'gray85') +
  geom_line(aes(year, Q), ldsRec, colour = 'steelblue') +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  facet_wrap(vars(ID), ncol = 1, scales = 'free_y', 
             strip.position = 'right') +
  labs(x = NULL, y = 'Q [m\u00b3/s]')
```

